sWGS validation report

### 2019-08-20 Validation Report.Rmd
...

@@

sheets = readxl::excel_sheets(val.file)[1:12]
all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
}))

@@

sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, Cohort) %>% dplyr::summarise_at(vars(varMAD_median), funs(min,mean,median,max),na.rm=T)

@@

@@

val %>% group_by(as.integer(RA),`Sample Type`, Collection) %>% dplyr::summarise(
  `Total Samples` = length(Batch),
  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
) %>% rename('as.integer(RA)' = 'Processing RA') %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')

@@

val %>% ungroup %>% filter(Batch <= 2) %>% dplyr::group_by(Batch, `Sample Type`) %>% dplyr::summarise(
  `Total Samples` = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

@@

```{r}
val %>% ungroup %>% filter(Batch %in% c(3,4)) %>% dplyr::group_by(Batch, `Sample Type`, Collection) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable(caption='Total samples and QC results') %>% kable_styling('striped')
```

@@

speedvac %>% dplyr::group_by(Batch, `Sample Type`, `Original Cohort`, Speedvac) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

@@

kit %>% dplyr::group_by(Batch, `Kit Extraction`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

@@

ci %>% dplyr::group_by(Batch, `Original cohort`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')
```

@@

on = left_join(val %>% ungroup %>% filter(Batch %in% c(8)),
               readxl::read_xlsx(val.file, sheet='Batch 8') %>% dplyr::select(`Block ID`, `Sample Date`, `Sample Type`))

@@

on = on %>% mutate( Age = ifelse( `Sample Date` < m18, '<18m', '>=18m') )

on %>% dplyr::group_by(Batch, `Sample Type`, `Age`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

if (file.exists('batch8_preds.txt')) {
  batch8_preds = read_tsv('batch8_preds.txt')
} else {
  segFiles = grep(paste(on$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)

  batch8_preds = do.call(bind_rows, lapply(segFiles, function(f) {
    print(f)
    load(f)
    if (nrow(sampleResiduals(segmented)) <= 0) return()

    segmented$sample.info = loadSampleInformation(  
      on %>% filter(sample %in% segmented$sample.info$Sample) %>%
        dplyr::rename('Endoscopy' = `Sample Date`, Sample = sample ) )

    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, F)
    BarrettsProgressionRisk::predictions(prr,'sample')
  }))
  batch8_preds = batch8_preds %>% filter(!is.na(`Patient ID`))

  write_tsv(batch8_preds, 'batch8_preds.txt')
}

batch8_preds %>% dplyr::select(`Patient ID`, Sample, everything(), -matches('Hospital|SLX|Index') ) %>% kable() %>% kable_styling('striped')

@@

oldp = val %>% ungroup %>% filter(Batch %in% c(8:12) & Cohort == 'Old progressor scrolls')

oldp %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
  `n Samples` = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

@@

sheets = readxl::excel_sheets(val.file)[9:12]
info = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
    dplyr::filter(Status == 'P') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% dplyr::mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
}))

if (file.exists('oldp_preds.txt')) {
  oldp.preds = read_tsv('oldp_preds.txt')
} else {
  segFiles = grep(paste(oldp$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)

#  info = readxl::read_xlsx(val.file, 'Batch 9') %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))

  oldp.preds = do.call(bind_rows, lapply(segFiles, function(f) {
    load(f)  
    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )

    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
    predictions(prr)
  }))
  write_tsv(oldp.preds, 'oldp_preds.txt')
}

oldp.preds = left_join(oldp.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))

oldp.preds = oldp.preds %>% filter(Sample %in% (oldp %>% filter(!grepl('OCCAMS',`Hospital Research ID`)) %>% dplyr::select(sample) %>% pull))

oldp.preds %>% dplyr::summarise(
  'TPR'=length(which(Risk == 'High'))/n(),
  'FNR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)

oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
  kable(caption='Count sample predicted risks, P') %>% kable_styling(full_width = F)

fnr.patients = oldp.preds %>% 
  group_by(`Hospital Research ID`) %>% dplyr::summarise(
  TPR = length(which(Risk == 'High'))/n()
) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull

oldp.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low')),
  'Pathology Notes' = paste(`Path Notes`, collapse='; ')
) %>% kable(caption = 'Classified predictions per patient (all P)') %>% kable_styling('striped')

@@

new.np = val %>% ungroup %>% filter(Batch %in% c(10:12) & Status == 'NP')

new.np %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
  Total = n(),
  `% Passed QC` = round((length(which(varMAD_median<=qc.cutoff))/n())*100,1)
) %>% kable() %>% kable_styling('striped')

@@

```{r}
sheets = readxl::excel_sheets(val.file)[10:12]
info = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s, trim_ws = T) %>% 
    dplyr::filter(Status == 'NP' & !is.na(`SLX-ID`)) %>% dplyr::select(`Hospital Research ID`, `Block ID`, Endoscopy, Pathology, `SLX-ID`, `Index Sequence`, `Path Notes`) %>% mutate(Sample = paste0(`SLX-ID`,'.',`Index Sequence`))
}))

if (file.exists('val_np.txt')) {
  new.np.preds = read_tsv('val_np.txt')
} else {
  segFiles = grep(paste(new.np$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '(2|3)_segObj',  full.names = T, recursive = T), value=T)

  new.np.preds = do.call(bind_rows, lapply(segFiles, function(f) {
    print(f)
    load(f)  
    if (nrow(BarrettsProgressionRisk::sampleResiduals(segmented) %>% dplyr::filter(Pass)) <= 0) return()

    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
    predictions(prr)
  }))
  write_tsv(new.np.preds, 'val_np.txt')
}

new.np.preds = left_join(new.np.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))

#cor.test(new.np.preds$n.segs, new.np.preds$varMAD_median)

new.np.preds %>% dplyr::summarise(
  'FPR'=length(which(Risk == 'High'))/n(),
  'TPR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)

new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
  kable(caption='Count sample predicted risks, NP') %>% kable_styling(full_width = F)

fpr.patients = new.np.preds %>%  group_by(`Hospital Research ID`) %>% dplyr::summarise(
  TPR = length(which(Risk == 'Low'))/n()
) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull

@@

with(val.new.np.preds, cor.test(varMAD_median, Probability))

val.new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low')),
  'QC var' = signif( mean(varMAD_median), 2)
) %>% kable(caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped')

@@

>>>>

pred.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/predictions'

@@

sheets = readxl::excel_sheets(val.file)[1:13]
all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
}))

@@

sval = val %>% group_by(Batch) %>% dplyr::mutate(Cohort = paste(unique(Cohort), collapse=', ' )) %>% group_by(Batch, Cohort) %>% 
  dplyr::summarise_at(vars(varMAD_median), list(~ min(., na.rm=T), ~ mean(., na.rm=T) ,~ median(., na.rm=T),~ max(., na.rm=T)))

@@

val %>% dplyr::group_by(Batch) %>% dplyr::summarise( Pass = length(which(varMAD_median <= qc.cutoff)), Fail = length(which(varMAD_median > qc.cutoff))  ) %>% kable(caption = 'Number of samples per batch that pass/fail') %>% kable_styling('striped', full_width = F)

@@

val %>% group_by(as.integer(RA),`Sample Type`, Collection) %>% dplyr::summarise(
  `Total Samples` = length(Batch),
  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
) %>% dplyr::rename('Processing RA' = 'as.integer(RA)' ) %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('striped')

@@

val %>% ungroup %>% filter(Batch <= 2) %>% dplyr::group_by(Batch, `Sample Type`) %>% dplyr::summarise(
  `Total Samples` = n(),
  `Passed QC` = length(which(varMAD_median<=qc.cutoff))
) %>% kable() %>% kable_styling('striped')

@@

```{r}
val %>% ungroup %>% filter(Batch %in% c(3,4)) %>% dplyr::group_by(Batch, `Sample Type`, Collection) %>% dplyr::summarise(
  Total = n(),
  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
) %>% kable(caption='Total samples and QC results') %>% kable_styling('striped')
```

@@

speedvac %>% dplyr::group_by(Batch, `Sample Type`, `Original Cohort`, Speedvac) %>% dplyr::summarise(
  Total = n(),
  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
) %>% kable() %>% kable_styling('striped')

@@

kit %>% dplyr::group_by(Batch, `Kit Extraction`) %>% dplyr::summarise(
  Total = n(),
  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
) %>% kable() %>% kable_styling('striped')

@@

ci %>% dplyr::group_by(Batch, `Original cohort`) %>% dplyr::summarise(
  Total = n(),
  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
) %>% kable() %>% kable_styling('striped')
```

@@

on = left_join(val %>% ungroup %>% filter(Batch %in% c(8)),
               readxl::read_xlsx(val.file, sheet='Batch 8') %>% dplyr::select(`Block ID`, `Endoscopy`, `Sample Type`))

@@

on = on %>% mutate( Age = ifelse( Endoscopy < m18, '<18m', '>=18m') )

on %>% dplyr::group_by(Batch, `Sample Type`, `Age`) %>% dplyr::summarise(
  Total = n(),
  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
) %>% kable() %>% kable_styling('striped')

val.info = do.call(bind_rows, lapply(sheets[8:13], function(s) {
  readxl::read_xlsx(val.file, s) %>% dplyr::select(-matches('Extraction|Qbit'))%>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% fncols('Collection', 'Biopsy')
    #dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) 
})) %>% mutate(
  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
  Samplename = paste(`SLX-ID`,`Index Sequence`,sep='.')
  )

preds.dir = paste0(analysis.dir, '/predictions/0.9')

batch8_preds = do.call(bind_rows,
  purrr::map(grep( paste(unique(on$`Patient ID`), collapse='|'), list.files(preds.dir, full.names = T, recursive = T), value=T), function(f) {
             read_tsv(f, col_types = cols(
                Sample = col_character(),
                Probability = col_double(),
                `Relative Risk` = col_double(),
                Risk = col_character(),
                `Hospital Research ID` = col_character(),
                `Block ID` = col_character(),
                Endoscopy = col_date(format = ""),
                Pathology = col_character(),
                `SLX-ID` = col_character(),
                `Index Sequence` = col_character(),
                `Path Notes` = col_logical()
      ))
})) %>% dplyr::mutate_if(is.numeric, list(round), 2)

batch8_preds %>% dplyr::select(`Hospital Research ID`, Sample, everything(), -matches('SLX|Index') ) %>% kable() %>% kable_styling('striped')

@@

oldp = val %>% ungroup %>% filter(Batch %in% c(8:13) & Cohort == 'Old progressor scrolls')

oldp %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
  `n Samples` = n(),
`Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
) %>% kable() %>% kable_styling('striped')

@@

The predictions from these samples are not particularly good.

```{r}
p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))

oldp.preds = do.call(bind_rows,
  purrr::map(grep( paste(unique(oldp$`Patient ID`), collapse='|'), list.files(preds.dir, full.names = T, recursive = T), value=T), function(f) {
             read_tsv(f,col_types = cols(
                Sample = col_character(),
                Probability = col_double(),
                `Relative Risk` = col_double(),
                Risk = col_character(),
                `Hospital Research ID` = col_character(),
                `Block ID` = col_character(),
                Endoscopy = col_date(format = ""),
                Pathology = col_character(),
                `SLX-ID` = col_character(),
                `Index Sequence` = col_character(),
                `Path Notes` = col_character()
      ))
})) %>% dplyr::mutate_if(is.numeric, list(round), 2) %>% dplyr::select(-Endoscopy)

oldp.preds = left_join(oldp.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))
oldp.preds = left_join(oldp.preds, val.info %>% dplyr::select(Samplename, Endoscopy), by=c('Sample'='Samplename'))
oldp.preds = oldp.preds %>% filter(Sample %in% (oldp %>% filter(!grepl('OCCAMS',`Hospital Research ID`)) %>% dplyr::select(sample) %>% pull))

oldp.preds = left_join(oldp.preds, p.diag, by=c('Hospital Research ID', 'Block ID', 'Endoscopy' = 'Sequenced Endoscopy'))

oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>%
  kable(caption='Progressor sample predicted risks') %>% kable_styling(full_width = F)

ggplot(oldp.preds, aes(`Hospital Research ID`, Probability, color=Probability)) + geom_point(size=2) +
  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
```

```{r}
oldp.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low')),
  'Pathology Notes' = paste(unique(`Path Notes`), collapse='; ')
) %>% dplyr::arrange(desc(`High Risk`), desc(`Mod. Risk`), desc(`Low Risk`)) %>% 
  mutate(
  `High Risk` = cell_spec(`High Risk`, "html", color = ifelse(`High Risk` > 0, "red", "grey39")),
  `Mod. Risk` = cell_spec(`Mod. Risk`, "html", color = ifelse(`Mod. Risk` > 0, "orange", "grey39")),
  `Low Risk` = cell_spec(`Low Risk`, "html", color = ifelse(`Low Risk` > 0, "blue", "grey39")),
) %>% kable(format='html',escape = F,caption = 'Classified predictions per patient (all Progressors)') %>% kable_styling('striped', full_width = F)

oldp.preds = oldp.preds %>% separate(`Block ID`, c('PID','Block'), '\\s?(-|\\s)', remove = F) 
```

However, it looks very likely that the age of the sample is related to the predictability. Roughly speaking, samples over 3.5 years old appear more likely to predict as low risk.  This wasn't so true of Ellie's samples.  These samples did sit as parafiin scrolls for a year after being cut as well and this could have affected the samples.

```{r}
ggplot(oldp.preds, aes(`Weeks Follow Up`, Probability, color=Probability)) + geom_point(size=2) +
  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + labs(title='Progressors', subtitle='Predictions by follow up time')

oldp.per.endo = 
  oldp.preds %>% group_by(`Hospital Research ID`, PID) %>% 
  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
  dplyr::summarise_at(vars(matches('Prob|Relative|Endo|n.samples')), list(mean)) %>% 
  dplyr::select(-matches('^Sample$')) %>% 
  arrange(`Hospital Research ID`, Probability, Endoscopy)

fnr.patients = oldp.preds %>% 
  group_by(`Hospital Research ID`) %>% dplyr::summarise(
  FNR = length(which(Risk == 'Low'))/n() ) %>% 
  dplyr::summarise( 'Misclassified patients' = length(which(FNR > 0.5))/n() ) %>% pull %>% round(2)

@@

```{r}
new.np = val %>% ungroup %>% filter(Batch %in% c(10:13) & Status == 'NP')

new.np.preds = do.call(bind_rows,
  purrr::map(grep( paste(unique(new.np$`Patient ID`), collapse='|'), list.files(preds.dir, full.names = T, recursive = T), value=T), function(f) {
             read_tsv(f,col_types = cols(
                Sample = col_character(),
                Probability = col_double(),
                `Relative Risk` = col_double(),
                Risk = col_character(),
                `Hospital Research ID` = col_character(),
                `Block ID` = col_character(),
                Endoscopy = col_date(format = ""),
                Pathology = col_character(),
                `SLX-ID` = col_character(),
                `Index Sequence` = col_character(),
                `Path Notes` = col_character()
      ))
})) %>% dplyr::mutate_if(is.numeric, list(round), 2) %>% dplyr::select(-Endoscopy)

new.np.preds = left_join(new.np.preds, val %>% dplyr::select(sample, varMAD_median, n.segs), by=c('Sample'='sample'))
new.np.preds = left_join(new.np.preds, val.info %>% dplyr::select(Samplename, Endoscopy), by=c('Sample'='Samplename'))

new.np %>% dplyr::group_by(Batch, `Sample Type`, `Cohort`) %>% dplyr::summarise(
  Total = n(),
  `Passed QC` = paste0(length(which(varMAD_median<=qc.cutoff)), ' (', round((length(which(varMAD_median<=qc.cutoff))/n())*100,1), '%)')
) %>% kable() %>% kable_styling('striped')

@@

new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>% 
  kable(caption='Count sample predicted risks, NP') %>% kable_styling(full_width = F)

@@

new.np.preds = new.np.preds %>% dplyr::mutate( 'Weeks Since Endoscopy' = as.double(difftime( as.Date('2019/06/01'), Endoscopy, units='weeks' )))

ggplot(new.np.preds, aes(`Hospital Research ID`, Probability, color=Probability)) + geom_point(size=2) +
  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))
```

Time since the endoscopy was performed doesn't appear to have any effect on the non-progressor predictions

```{r}
ggplot(new.np.preds, aes(`Weeks Since Endoscopy`, Probability, color=Probability)) + geom_point(size=2) +
  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + labs(title='Non-progressors', subtitle='Predictions vs time since the endoscopy')

val.new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(),
  'High Risk' =  length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low')),
  'QC var' = signif( mean(varMAD_median), 2)
) %>% dplyr::arrange( `Low Risk`, `Mod. Risk`, `High Risk`) %>% 
  mutate(
  `High Risk` = cell_spec(`High Risk`, "html", color = ifelse(`High Risk` > 0, "red", "grey39")),
  `Mod. Risk` = cell_spec(`Mod. Risk`, "html", color = ifelse(`Mod. Risk` > 0, "orange", "grey39")),
  `Low Risk` = cell_spec(`Low Risk`, "html", color = ifelse(`Low Risk` > 0, "blue", "grey39")),
) %>%
  kable(format='html',escape = F, caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped', full_width = F)

new.np.preds = new.np.preds %>% separate(`Block ID`, c('PID','Block'), '\\s?(-|\\s)', remove = F)

np.per.endo = 
  new.np.preds %>% group_by(`Hospital Research ID`, PID) %>% 
  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
  dplyr::summarise_at(vars(matches('Prob|Relative|Endo|n.samples')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
  arrange(`Hospital Research ID`, Probability, PID)

@@

## All together

The predictions show a very poor AUC as the split shown above would make you expect.

all = bind_rows(
  oldp.preds %>% dplyr::select(Sample, Probability, `Relative Risk`, Risk) %>% dplyr::mutate(Status = 'P'),
  new.np.preds %>% dplyr::select(Sample, Probability, `Relative Risk`, Risk) %>% dplyr::mutate(Status = 'NP')
)

roc = pROC::roc(Status ~ Probability, data = all, auc=T, ci=T, of='thresholds',transpose=T)
roc.plot(roc)
#pROC::coords(roc, 'best', transpose=T)
```

```{r, fig.height=8}

p1 = ggplot(all, aes(Probability)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + 
  scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
  plot.theme + labs(title='Sample probability', y='n Samples', x='Probability') + theme(legend.position = 'none')

p2 = ggplot(all, aes(`Relative Risk`)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme

grid.arrange(p1,p2)
